---
title: 'Deep Learning'
author: "Laxman Panthi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# load r libraries
library(readr)
library(reticulate)
use_condaenv('anly540', required = T)
py_config()
```


## The Model

Since the deep models take a long time to run and often require a lot of computer power, we will just detail where you might use a deep learning model in your own work. 

In this section, detail a classification task you might use a deep learning model for:

### Load data
Data was downloaded from kaggle. It has two columns:
- Text: Twitter text from multiple twitter handles
- Sentiment: Classification if the text was positive or negative.
```{r}
data <- read_csv("stock_sentiments.csv")
head(data)
```

```{python}
data_py = r.data

# check how it looks in python
data_py.head()
```


  - What are the class labels? (i.e. what are the groups?)

The field Sentiment is the dependent variable with two classes 1 & -1 where 1 means positive sentiment and -1 means negative sentiment.

```{python}
# lets look at the value counts of the classification
data_py.Sentiment.value_counts()
```

  - What would you use to predict the labels? Explain how text is involved in the prediction.

The predictor is the text or tweet from twitter users. 
We will have to clean the data - remove punctuations, remove stop words, lowercase etc.
Then the data is tokenized to create a new feature 'Token' using word2vec or similar model.
The data is then split to train & test (70-30 split)

  - What type of model would you use (RNN, CNN, mix of both)?

We will be using CNN in this case as the data will be in form of embedding matrix.

  - How many layers would you include in your model?

There will be six layers:
Embedding layer
DropOut
Conv1D
Maxpooling
Dense
Output using relu activation since it's a binary classfication model.

  - What type of layers do you think might be effective to help you predict your classes? 

We could add one more dropout and dense layer but it would require some testing. A simple model is usually better but if the data is complex we could add more layers.
  


