---
title: 'Word2Vec'
author: "Laxman Panthi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries + functions

Load all the libraries or functions that you will use to for the rest of the assignment. It is helpful to define your libraries and functions at the top of a report, so that others can know what they need for the report to compile correctly.

```{r libaries}
##r chunk
library(tidyverse)
library(reticulate)
use_condaenv("anly540")
reticulate::py_config()
```

Load the Python libraries or functions that you will use for that section. 

```{python}
##python chunk
import nltk
import gensim
from nltk.corpus import abc
import pandas as pd
import numpy as np
from bs4 import BeautifulSoup
import re
from nltk.corpus import stopwords
from sklearn.metrics import classification_report
from sklearn.metrics import accuracy_score
```

## The Data

The dataset is a set of Youtube comments that have been coded as:
  - 1: spam youtube messages 
  - 0: good youtube messages
  - This data is stored in the `CLASS` column

Import the data using either R or Python. I put a Python chunk here because you will need one to import the data, but if you want to first import into R, that's fine. 

### Load data in R
```{r}
##r chunk
data <- read_csv("youtube_spam.csv")
summary(data)
head(data)
```


### Move from r to python
```{python}
##python chunk
data_py = r.data
data_py.head
```

## Clean up the data

Use one of our clean text functions to clean up the `CONTENT` column in the dataset. 

```{python}
##python chunk
REPLACE_BY_SPACE_RE = re.compile('[/(){}\[\]\|@,;]') #remove symbols with space
BAD_SYMBOLS_RE = re.compile('[^0-9a-z #+_]') #take out symbols altogether
STOPWORDS = set(stopwords.words('english')) #stopwords

def clean_text(text):
    text = BeautifulSoup(text, "lxml").text # HTML decoding
    text = text.lower() # lowercase text
    text = REPLACE_BY_SPACE_RE.sub(' ', text) # replace REPLACE_BY_SPACE_RE symbols by space in text
    text = BAD_SYMBOLS_RE.sub('', text) # delete symbols which are in BAD_SYMBOLS_RE from text
    text = ' '.join(word for word in text.split() if word not in STOPWORDS) # delete stopwors from text
    return text
    
data_py['CONTENT'] = data_py['CONTENT'].apply(clean_text)
data_py.head()
```

## Split the data

Split the data into testing and training data.

```{python}
##python chunk
##python chunk
X = data_py['CONTENT']
y = data_py['CLASS']

from sklearn.model_selection import train_test_split
X_train, X_test, y_train, y_test = train_test_split(X, y, test_size=0.20, random_state = 42)

```

## Process the data

For FastText and word2vec, create the tokenized vectors of the text.

```{python}
##python chunk
tokenized_train = [nltk.tokenize.word_tokenize(text)
                   for text in X_train.to_list()]
tokenized_test = [nltk.tokenize.word_tokenize(text)
                   for text in X_test.to_list()]
```

## Word2Vec

Build the word2vec model.

```{python}
##python chunk
# build word2vec model
w2v_model = gensim.models.Word2Vec(tokenized_train, 
                                   size=100, window=6,
                                   min_count=2, iter=5, workers=4)
```

## Convert the model

Convert the model into a set of features to use in our classifier. 

```{python}
##python chunk
def document_vectorizer(corpus, model, num_features):
    vocabulary = set(model.wv.index2word)
    
    def average_word_vectors(words, model, vocabulary, num_features):
        feature_vector = np.zeros((num_features,), dtype="float64")
        nwords = 0.
        
        for word in words:
            if word in vocabulary: 
                nwords = nwords + 1.
                feature_vector = np.add(feature_vector, model.wv[word])
        if nwords:
            feature_vector = np.divide(feature_vector, nwords)

        return feature_vector

    features = [average_word_vectors(tokenized_sentence, model, vocabulary, num_features)
                    for tokenized_sentence in corpus]
    return np.array(features)
    
# generate averaged word vector features from word2vec model
avg_wv_train_features = document_vectorizer(corpus=tokenized_train,
                                                    model=w2v_model,
                                                     num_features=100)
avg_wv_test_features = document_vectorizer(corpus=tokenized_test, 
                                                    model=w2v_model,
                                                    num_features=100)
#define your outcomes
my_tags = ['0', '1']
```

## Build a classifier model

In class, we used logistic regression to classify the data. You can use any machine learning algorithm you want here, and build a classification model. 

### Logistic Regression

```{python}
##python chunk
#build a log model
from sklearn.linear_model import LogisticRegression
logreg = LogisticRegression(solver='lbfgs', multi_class='ovr', max_iter=10000)

#fit the data to the log model
logreg = logreg.fit(avg_wv_train_features, y_train)
```

#### Examine the results

Print out the accuracy, recall, and precision of your model.

```{python}
##python chunk
#predict new data
y_pred = logreg.predict(avg_wv_test_features)

#print out results
print('accuracy %s' % accuracy_score(y_pred, y_test))
print(classification_report(y_test, y_pred,target_names=my_tags))
```

### Support Vector Machines

```{python}
##python chunk
#build a svm model
from sklearn.linear_model import SGDClassifier

sgd = Pipeline([('vect', CountVectorizer()),
                ('tfidf', TfidfTransformer()),
                ('clf', SGDClassifier(loss='hinge', penalty='l2',alpha=1e-3, random_state=42, max_iter=5, tol=None)),
               ])
sgd.fit(X_train, y_train)

%%time

y_pred = sgd.predict(X_test)

print('accuracy %s' % accuracy_score(y_pred, y_test))
print(classification_report(y_test, y_pred,target_names=my_tags))
```

#### Examine the results

Print out the accuracy, recall, and precision of your model.

```{python}
##python chunk
#predict new data
y_pred = logreg.predict(avg_wv_test_features)

#print out results
print('accuracy %s' % accuracy_score(y_pred, y_test))
print(classification_report(y_test, y_pred,target_names=my_tags))
```

## Build a FastText model

Using the same data, build a FastText model.

```{python}
##python chunk
##python chunk
from gensim.models.fasttext import FastText

#build a fast test model
ft_model = FastText(tokenized_train, size=100, window=6, 
                    min_count=2, iter=5, workers=4)
```

## Extract the features

Convert the FastText model into features for prediction.

```{python}
##python chunk
avg_ft_train_features = document_vectorizer(corpus=tokenized_train, model=ft_model,
                                                     num_features=100)
avg_ft_test_features = document_vectorizer(corpus=tokenized_test, model=ft_model,
                                                    num_features=100)           
```

## Build a classifier model

Using the same machine learning algorithm as above, build a classifier model that uses the FastText data to predict the categories.

### Logistic Regression

```{python}
##python chunk
from sklearn.linear_model import LogisticRegression
logreg = LogisticRegression(solver='lbfgs', multi_class='ovr', max_iter=10000)
logreg = logreg.fit(avg_ft_train_features, y_train)
y_pred = logreg.predict(avg_ft_test_features)
```

#### Examine the results

Print out the accuracy, recall, and precision of your model.

```{python}
##python chunk
print('accuracy %s' % accuracy_score(y_pred, y_test))
print(classification_report(y_test, y_pred,target_names=my_tags))
```

## Interpretation 

- Where you able to predict the spam messages from the real comments? 
- Which model provided you with a better prediction?

Yes. The prediction model was created with 63% accuracy. Both the model had similar accuracy, precision and recall.